---
title: "Data Manipulation"
author: "Xue Yang"
date: "9/20/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)   # collection of some package which make things more easily
```


## Import FAS csv files

Import my csv files to use as examples.

```{r}
litters_data = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data/FAS_pups.csv",
  col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)
```


## Select variable

... by listing

```{r}
# ?select
# View(litters_data)

select(litters_data, group, litter_number, gd0_weight)

select(litters_data, litter_number, gd0_weight, pups_survive)
```

... by sepecifying a range:


```{r}
select(litters_data, gd_of_birth:pups_survive)
select(litters_data, group, gd_of_birth:pups_survive)
```


... by saying what you want to remove

```{r}
select(litters_data, -litter_number)
```

... and rename

```{r}
select(litters_data, group, litter_number, gest_day_0_weight = gd0_weight)
rename(litters_data, gest_day_0_weight = gd0_weight)



```


... use select_helpers

```{r}

# ?select_helpers

select(litters_data, starts_with("gd"))

select(litters_data, litter_number, starts_with("pup"))

# first select litter_number and then the rest of everything
select(litters_data, litter_number, everything())
```

... Assessment:In the pups data, select the columns containing litter number, sex, and PD ears.
```{r}
select(pups_data, litter_number, sex, pd_ears)
```


## Filter observation

Do some filtering!!

```{r}
filter(litters_data, gd0_weight < 25)
filter(litters_data, gd0_weight >= 25)

filter(litters_data, pups_born_alive == 8)

filter(litters_data, is.na(gd0_weight))
filter(litters_data, !is.na(gd0_weight))

filter(litters_data, group == "Low7")

filter(litters_data, group == "Low8" | group == "Low7")

# use %in% to detect if values appear in a set
filter(litters_data, group %in% c("Low8",  "Low7"))

filter(litters_data, pups_born_alive %in% 4:6)

filter(litters_data, pups_born_alive %in% 4:7, !is.na(gd0_weight))
``` 

for pup_data: filter to include only pups with sex 1, filter to include only pups with PD walk less than 11 and sex 2

```{r}
filter(pups_data, sex == 1)

filter(pups_data, sex == 2, pd_walk < 11)
```



## Mutate

Creat variables

```{r}
mutate(litters_data, 
       wt_gain = gd18_weight - gd0_weight,
       group = tolower(group))

mutate(litters_data,
       wt_gain = gd18_weight - gd0_weight,
       wt_gain_squared = wt_gain^2)
```


## Arrange

Arrange the data...

```{r}
arrange(litters_data, gd0_weight)

arrange(litters_data, pups_born_alive, gd0_weight)
arrange(litters_data,desc(pups_born_alive), gd0_weight)
```

## Piping

Looking at 

```{r}
litters_data_raw = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data_clean_names = janitor::clean_names(litters_data_raw)
litters_data_selected_cols = select(litters_data_clean_names, -pups_survive)
litters_data_with_vars = mutate(litters_data_selected_cols, 
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group))
litters_data_with_vars
```


```{r}
litters_data_clean = 
  mutate(
    select(
      janitor::clean_names(
        read_csv("./data/FAS_litters.csv", col_types = "ccddiiii")
        ), 
    -pups_survive
    ),
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group)
  )
litters_data_clean
```


using piping

```{r}
read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = tolower(group)
  ) %>%
  filter(.data = .,!is.na(gd0_weight))
```



Illustrate placeholder use with 'lm'

```{r}
read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
     wt_gain = gd18_weight - gd0_weight,
    group = tolower(group)
  ) %>%
  filter(!is.na(gd0_weight)) %>%
  lm(gd18_weight ~ gd0_weight, data = .)

```


